---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
params:
  seurat_obj: NULL
  comp_group: "comparison"
  author: "Pipeline"
  title: "QC Report"
  out_dir: "."
  cfg: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(Seurat)
library(ggplot2)
library(dplyr)
library(knitr)
```

```{r load-obj}
obj <- params$seurat_obj
if (is.null(obj)) stop("No Seurat object provided")
group  <- params$comp_group
out_dir <- params$out_dir
cfg    <- params$cfg   # may be NULL when report is rendered standalone
```

# QC Report: `r group`

## Summary of the dataset

```{r summary}
total_cells       <- ncol(obj)
total_genes       <- nrow(obj)
cells_by_sample   <- table(obj$orig.ident)
cells_by_condition <- table(obj$condition)
```

| Metric                | Value                                          |
|-----------------------|------------------------------------------------|
| **Total cells**       | `r total_cells`                                |
| **Total genes**       | `r total_genes`                                |
| **Samples**           | `r length(unique(obj$orig.ident))`             |
| **Conditions**        | `r length(unique(obj$condition))`              |

**Cells per sample:**
```{r cells-by-sample, results='asis'}
kable(as.data.frame(cells_by_sample), col.names = c("Sample", "Number of cells"))
```

**Cells per condition:**
```{r cells-by-condition, results='asis'}
kable(as.data.frame(cells_by_condition), col.names = c("Condition", "Number of cells"))
```

## Filtering history

```{r filtering-params, results='asis'}
# cfg is forwarded from the pipeline; show thresholds when available.
if (!is.null(cfg)) {
  params_df <- data.frame(
    Parameter = c(
      "Min features", "Max features", "Max counts",
      "Max mitochondrial %", "Gene removal pattern"
    ),
    Value = c(
      cfg$qc$min_features,
      cfg$qc$max_features  %||% "N/A",
      cfg$qc$max_counts,
      cfg$qc$max_mt_percent,
      cfg$species$gene_removal_pattern
    ),
    stringsAsFactors = FALSE
  )
  kable(params_df)
} else {
  cat("_(Config not available – run via `sparks()` to populate this section.)_\n")
}
```

Doublet removal used scDblFinder. Per-sample stats are in `QC/Doublet_Removed_*.txt`.

## QC metrics after filtering

```{r qc-violin, fig.width=12, fig.height=5}
p1 <- VlnPlot(obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),
              ncol = 3, pt.size = 0.1)
print(p1)
```

```{r qc-scatter, fig.width=12, fig.height=5}
p2 <- FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "percent.mt")
p3 <- FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
print(p2 + p3)
```

## Doublet detection

```{r doublets}
if ("scDblFinder.class" %in% colnames(obj[[]])) {
  dtab <- table(obj$scDblFinder.class)
  kable(as.data.frame(dtab), col.names = c("Class", "Number of cells"))
} else {
  cat("No doublet classification found in the merged object.",
      "(Doublets were removed before merging; see per-sample stats in the QC folder.)\n")
}
```

## Cell cycle and sex scoring

```{r scoring}
if ("S.Score" %in% colnames(obj[[]])) {
  cat("**Cell cycle phase distribution**\n")
  print(kable(as.data.frame(table(Phase = obj$Phase)),
              col.names = c("Phase", "Cells")))
}
if ("Assigned_Sex" %in% colnames(obj[[]])) {
  cat("\n**Sex assignment**\n")
  print(kable(as.data.frame(table(Sex = obj$Assigned_Sex)),
              col.names = c("Sex", "Cells")))
}
if (!"S.Score" %in% colnames(obj[[]]) && !"Assigned_Sex" %in% colnames(obj[[]])) {
  cat("Cell cycle and sex scoring were not run for this dataset.\n")
}
```

## Clustering overview

Number of clusters: `r length(unique(obj$seurat_clusters))`

```{r umap, fig.width=8, fig.height=6}
DimPlot(obj, reduction = "umap", group.by = "seurat_clusters", label = TRUE) +
  ggtitle(paste("UMAP -", group))
```

## Session info

```{r session}
sessionInfo()
```
