---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
params:
  seurat_obj: NULL
  comp_group: "comparison"
  author: "Pipeline"
  title: "QC Report"
  out_dir: "."
  cfg: NULL
  run_files: NULL
  log_path: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(Seurat)
library(ggplot2)
library(dplyr)
library(knitr)

# `%||%` is exported by SPARKS but is not on the search path when the package is
# only loaded (SPARKS::sparks()), and base R only has it from 4.4.0.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(lhs, rhs) if (is.null(lhs) || length(lhs) == 0L) rhs else lhs
}
```

```{r load-obj}
obj <- params$seurat_obj
if (is.null(obj)) stop("No Seurat object provided")
group  <- params$comp_group
out_dir <- params$out_dir
cfg    <- params$cfg   # may be NULL when report is rendered standalone
```

# QC Report: `r group`

## Summary of the dataset

```{r summary}
total_cells <- ncol(obj)
total_genes <- nrow(obj)
cells_by_sample <- if ("orig.ident" %in% colnames(obj[[]])) table(obj$orig.ident) else integer(0)
cells_by_condition <- if ("condition" %in% colnames(obj[[]])) table(obj$condition) else integer(0)
```

| Metric                | Value                                          |
|-----------------------|------------------------------------------------|
| **Total cells**       | `r total_cells`                                |
| **Total genes**       | `r total_genes`                                |
| **Samples**           | `r if ("orig.ident" %in% colnames(obj[[]])) length(unique(obj$orig.ident)) else "Not available"` |
| **Conditions**        | `r if ("condition" %in% colnames(obj[[]])) length(unique(obj$condition)) else "Not available"` |

**Cells per sample:**
```{r cells-by-sample, results='asis'}
if (length(cells_by_sample)) {
  print(kable(as.data.frame(cells_by_sample), col.names = c("Sample", "Number of cells")))
} else {
  cat("_(Sample metadata not available.)_\n")
}
```

**Cells per condition:**
```{r cells-by-condition, results='asis'}
if (length(cells_by_condition)) {
  print(kable(as.data.frame(cells_by_condition), col.names = c("Condition", "Number of cells")))
} else {
  cat("_(Condition metadata not available.)_\n")
}
```

## Input data

```{r input-data, results='asis'}
if ("quant_format" %in% colnames(obj[[]])) {
  fmt_df <- unique(obj@meta.data[, c("orig.ident", "quant_format"), drop = FALSE])
  colnames(fmt_df) <- c("Sample", "Format")
  # BUG FIX: a bare kable() inside an if-block is never auto-printed, so this
  # table silently never appeared in the report.
  print(kable(fmt_df))
  if (!is.null(cfg) && isTRUE(cfg$input$save_raw_matrix)) {
    cat("\nRaw, pre-QC count matrices were saved per sample under ",
        "`results_dir/", group, "/RawMatrix/<folder_id>/` in ",
        cfg$input$raw_matrix_format %||% "mtx", " format.\n", sep = "")
  }
} else {
  cat("_(Input format metadata not available for this run.)_\n")
}
```

## Filtering history

```{r filtering-params, results='asis'}
# cfg is forwarded from the pipeline; show thresholds when available.
if (!is.null(cfg)) {
  params_df <- data.frame(
    Parameter = c(
      "Min features", "Max features", "Max counts",
      "Max mitochondrial %", "Gene removal pattern"
    ),
    Value = c(
      cfg$qc$min_features,
      cfg$qc$max_features  %||% "N/A",
      cfg$qc$max_counts,
      cfg$qc$max_mt_percent,
      cfg$species$gene_removal_pattern
    ),
    stringsAsFactors = FALSE
  )
  print(kable(params_df))
} else {
  cat("_(Config not available – run via `sparks()` to populate this section.)_\n")
}
```

Doublet removal used scDblFinder. Per-sample stats are in `QC/Doublet_Removed_*.txt`.

## QC metrics after filtering

```{r qc-violin, fig.width=12, fig.height=5}
qc_features <- intersect(c("nFeature_RNA", "nCount_RNA", "percent.mt"),
                          colnames(obj[[]]))
if (length(qc_features)) {
  print(VlnPlot(obj, features = qc_features, ncol = min(3L, length(qc_features)),
                pt.size = 0.1))
} else {
  cat("QC metric columns are not available in this object.\n")
}
```

```{r qc-scatter, fig.width=12, fig.height=5}
scatter_features <- intersect(c("nCount_RNA", "percent.mt", "nFeature_RNA"),
                              colnames(obj[[]]))
if (all(c("nCount_RNA", "percent.mt", "nFeature_RNA") %in% scatter_features)) {
  p2 <- FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "percent.mt")
  p3 <- FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
  print(p2 + p3)
} else {
  cat("Scatter plot metadata is not available.\n")
}
```

## Doublet detection

```{r doublets}
if ("scDblFinder.class" %in% colnames(obj[[]])) {
  dtab <- table(obj$scDblFinder.class)
  print(kable(as.data.frame(dtab), col.names = c("Class", "Number of cells")))
} else {
  cat("No doublet classification found in the merged object.",
      "(Doublets were removed before merging; see per-sample stats in the QC folder.)\n")
}
```

## Cell cycle and sex scoring

```{r scoring}
if ("S.Score" %in% colnames(obj[[]])) {
  cat("**Cell cycle phase distribution**\n")
  print(kable(as.data.frame(table(Phase = obj$Phase)),
              col.names = c("Phase", "Cells")))
}
if ("Assigned_Sex" %in% colnames(obj[[]])) {
  cat("\n**Sex assignment**\n")
  print(kable(as.data.frame(table(Sex = obj$Assigned_Sex)),
              col.names = c("Sex", "Cells")))
}
if (!"S.Score" %in% colnames(obj[[]]) && !"Assigned_Sex" %in% colnames(obj[[]])) {
  cat("Cell cycle and sex scoring were not run for this dataset.\n")
}
```

## Clustering overview

Number of clusters: `r if ("seurat_clusters" %in% colnames(obj[[]])) length(unique(obj$seurat_clusters)) else "Not available"`

```{r umap, fig.width=8, fig.height=6}
if (!"seurat_clusters" %in% colnames(obj[[]])) {
  cat("Cluster metadata is not available.\n")
} else if (!"umap" %in% names(obj@reductions)) {
  cat("UMAP reduction is not available.\n")
} else {
  print(DimPlot(obj, reduction = "umap", group.by = "seurat_clusters", label = TRUE) +
          ggtitle(paste("UMAP -", group)))
}
```

## Pipeline diagnostics

```{r diagnostics, results='asis'}
if (!is.null(params$log_path) && file.exists(params$log_path)) {
  log_lines <- readLines(params$log_path, warn = FALSE)
  diagnostics <- grep("\\[(WARNING|ERROR)\\]", log_lines, value = TRUE)
  skips <- grep("\\[SKIP", log_lines, value = TRUE)
  if (length(diagnostics) || length(skips)) {
    diagnostic_df <- data.frame(
      Type = c(rep("Warning/error", length(diagnostics)),
               rep("Skipped/unavailable", length(skips))),
      Message = c(diagnostics, skips),
      stringsAsFactors = FALSE
    )
    print(kable(diagnostic_df))
  } else {
    cat("No warnings, errors, or skipped analyses were logged.")
  }
} else {
  cat("Pipeline log is not available.")
}
```

## Session info

```{r session}
sessionInfo()
```
