---
title:  "`r params$title`"
author: "`r params$author`"
date:   "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme:          flatly
    highlight:      tango
    toc:            true
    toc_float:
      collapsed:    false
      smooth_scroll: true
    toc_depth:      3
    number_sections: false
    self_contained: true    # embed all assets → single portable file
    code_folding:   hide
params:
  seurat_obj: NULL   # fully processed Seurat object
  comp_group: "Group"
  author:     "Pipeline User"
  title:      "SPARKS Results Report"
  out_dir:    "."
  cfg:        NULL
  groupings:  NULL   # character vector of metadata columns to iterate over
  run_files: NULL
  log_path: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# ── Packages ──────────────────────────────────────────────────────────────────
required_pkgs <- c("Seurat", "SeuratObject", "ggplot2", "dplyr",
                   "scales", "tidyr", "knitr")

missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace,
                                      quietly = TRUE, FUN.VALUE = logical(1))]
if (length(missing_pkgs) > 0L) {
  stop("Missing packages required for the results report: ",
       paste(missing_pkgs, collapse = ", "),
       "\nInstall them with install.packages(c(",
       paste0('"', missing_pkgs, '"', collapse = ", "), "))")
}

library(ggplot2)
library(dplyr)
has_plotly <- requireNamespace("plotly", quietly = TRUE)
has_dt <- requireNamespace("DT", quietly = TRUE)

# ── Helpers ───────────────────────────────────────────────────────────────────
# `%||%` is exported by SPARKS but is not on the search path when the package is
# only loaded (SPARKS::sparks()), and base R only has it from 4.4.0.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(lhs, rhs) if (is.null(lhs) || length(lhs) == 0L) rhs else lhs
}

.col_exists <- function(obj, col) col %in% colnames(obj@meta.data)

.nice_dt <- function(df, caption = NULL, page_len = 15, ...) {
  if (!has_dt) {
    return(knitr::kable(df, caption = caption, row.names = FALSE))
  }
  DT::datatable(
    df,
    caption    = caption,
    rownames   = FALSE,
    filter     = "top",
    extensions = c("Buttons", "Scroller"),
    options    = list(
      pageLength = page_len,
      scrollX    = TRUE,
      dom        = "Bfrtip",
      buttons    = c("copy", "csv", "excel"),
      ...
    ),
    class = "compact stripe hover"
  )
}

.umap_plotly <- function(obj, color_by, title = color_by,
                          point_size = 2, alpha = 0.7) {
  if (!color_by %in% colnames(obj@meta.data)) return(NULL)
  emb <- as.data.frame(Seurat::Embeddings(obj, "umap"))
  colnames(emb) <- c("UMAP_1", "UMAP_2")
  emb[[color_by]] <- as.character(obj@meta.data[[color_by]])
  emb$cell        <- rownames(emb)

  # Build a consistent discrete palette
  lvls <- unique(emb[[color_by]])
  n    <- length(lvls)
  pal  <- scales::hue_pal()(n)
  names(pal) <- lvls

  p <- ggplot(emb, aes(
      x    = UMAP_1, y = UMAP_2,
      color = .data[[color_by]],
      text  = paste0("Cell: ", cell, "<br>", color_by, ": ", .data[[color_by]])
    )) +
    geom_point(size = point_size, alpha = alpha) +
    scale_color_manual(values = pal) +
    labs(title = title, color = color_by) +
    theme_minimal(base_size = 11) +
    theme(legend.title = element_text(size = 9),
          legend.text  = element_text(size = 8),
          plot.title   = element_text(size = 12, face = "bold"))

  if (has_plotly) {
    plotly::ggplotly(p, tooltip = "text") |>
      plotly::layout(legend = list(itemsizing = "constant")) |>
      plotly::config(displayModeBar = TRUE, scrollZoom = TRUE)
  } else {
    p
  }
}

# ── Materialise params ────────────────────────────────────────────────────────
obj        <- params$seurat_obj
comp_group <- params$comp_group
cfg        <- params$cfg
out_dir    <- params$out_dir
groupings  <- params$groupings
run_files <- params$run_files

.current_files <- function(files) {
  if (length(files) == 0L || is.null(run_files)) return(files)
  files <- normalizePath(files, winslash = "/", mustWork = FALSE)
  intersect(files, normalizePath(run_files, winslash = "/", mustWork = FALSE))
}

# Derive groupings if not provided
if (is.null(groupings) || length(groupings) == 0L) {
  singler_cols <- grep("^singleR_labels", colnames(obj@meta.data), value = TRUE)
  groupings    <- unique(c("seurat_clusters", singler_cols))
  # BUG FIX: was vapply(groupings, .col_exists, obj, FUN.VALUE = ...), which
  # passes `obj` as the 2nd argument (`col`) and the grouping name as the 1st
  # (`obj`) -> "no applicable method for `@` applied to ... character".
  groupings    <- groupings[vapply(groupings, function(g) .col_exists(obj, g),
                                   FUN.VALUE = logical(1))]
}
```

---

## 1 · Dataset Overview {.tabset}

```{r overview}
n_cells      <- ncol(obj)
n_genes      <- nrow(obj)
n_clusters   <- if (.col_exists(obj, "seurat_clusters"))
                  length(unique(obj$seurat_clusters)) else NA
n_conditions <- if (.col_exists(obj, "condition"))
                  length(unique(obj$condition)) else NA
n_samples    <- if (.col_exists(obj, "orig.ident"))
                  length(unique(obj$orig.ident)) else NA

overview_df <- data.frame(
  Metric = c("Comparison group", "Cells (post-QC)",
             "Genes", "Clusters", "Conditions", "Samples"),
  Value  = c(comp_group,
             format(n_cells,      big.mark = ","),
             format(n_genes,      big.mark = ","),
             ifelse(is.na(n_clusters),   "—", n_clusters),
             ifelse(is.na(n_conditions), "—", n_conditions),
             ifelse(is.na(n_samples),    "—", n_samples))
)
knitr::kable(overview_df, col.names = c("Metric", "Value"),
             align = c("l", "r"))
```

```{r qc-params, results='asis'}
if (!is.null(cfg$qc)) {
  cat("\n### QC Thresholds Applied\n\n")
  qcp <- cfg$qc
  qc_df <- data.frame(
    Parameter = c("min_features", "max_features", "max_counts",
                  "max_mt_percent", "min_cells"),
    Value     = c(qcp$min_features %||% "—", qcp$max_features %||% "—",
                  qcp$max_counts   %||% "—", qcp$max_mt_percent %||% "—",
                  qcp$min_cells    %||% "—")
  )
  print(knitr::kable(qc_df, col.names = c("Parameter", "Value"), align = c("l","r")))
}
```

```{r input-data, results='asis'}
if (!is.null(cfg$input) && .col_exists(obj, "quant_format")) {
  cat("\n### Input Data\n\n")
  fmt_df <- unique(obj@meta.data[, c("orig.ident", "quant_format"), drop = FALSE])
  colnames(fmt_df) <- c("Sample", "Format")
  print(knitr::kable(fmt_df, align = c("l", "l")))
  if (isTRUE(cfg$input$save_raw_matrix)) {
    cat("\nRaw, pre-QC count matrices were saved per sample under ",
        "`results_dir/", comp_group, "/RawMatrix/<folder_id>/` in ",
        cfg$input$raw_matrix_format %||% "mtx", " format.\n", sep = "")
  }
}
```

---

## 2 · Quality Control {.tabset}

### Cell-level metrics

```{r qc-violin, fig.height=5, fig.width=11}
qc_meta_cols <- intersect(
  c("nFeature_RNA", "nCount_RNA", "percent.mt"),
  colnames(obj@meta.data)
)
group_col_for_vln <- if (.col_exists(obj, "orig.ident")) "orig.ident" else
                     if (.col_exists(obj, "condition"))  "condition"  else NULL

if (length(qc_meta_cols) > 0L && !is.null(group_col_for_vln)) {
  vln_data <- obj@meta.data[, c(qc_meta_cols, group_col_for_vln), drop = FALSE]
  vln_long <- tidyr::pivot_longer(
    vln_data,
    cols       = all_of(qc_meta_cols),
    names_to   = "Metric",
    values_to  = "Value"
  )
  p <- ggplot(vln_long,
              aes(x = .data[[group_col_for_vln]], y = Value,
                  fill = .data[[group_col_for_vln]])) +
    geom_violin(alpha = 0.7, scale = "width") +
    geom_boxplot(width = 0.1, outlier.size = 0.3, outlier.alpha = 0.3,
                 fill = "white", alpha = 0.6) +
    facet_wrap(~ Metric, scales = "free_y", ncol = 3) +
    scale_fill_brewer(palette = "Set2") +
    labs(x = NULL, y = NULL, fill = NULL) +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 35, hjust = 1),
          legend.position = "none")
  if (has_plotly) plotly::ggplotly(p) else p
} else {
  cat("> QC metric columns not found in metadata.")
}
```

### Per-sample cell counts

```{r qc-counts}
if (.col_exists(obj, "orig.ident")) {
  sample_counts <- as.data.frame(table(Sample = obj$orig.ident))
  colnames(sample_counts) <- c("Sample", "Cells")
  sample_counts$Cells <- format(sample_counts$Cells, big.mark = ",")
  .nice_dt(sample_counts, caption = "Cells per sample after QC filtering")
} else {
  cat("> orig.ident not found in metadata.")
}
```

---

## 3 · Clustering {.tabset}

### UMAP – Clusters

```{r umap-clusters, fig.height=6, fig.width=8}
.umap_plotly(obj, "seurat_clusters", title = paste("UMAP –", comp_group, "| Clusters"))
```

### UMAP – Condition

```{r umap-condition, fig.height=6, fig.width=8}
if (.col_exists(obj, "condition")) {
  .umap_plotly(obj, "condition", title = paste("UMAP –", comp_group, "| Condition"))
} else {
  cat("> 'condition' column not found.")
}
```

### Cluster composition table

```{r cluster-table}
if (.col_exists(obj, "seurat_clusters")) {
  clu_df <- as.data.frame(table(Cluster = obj$seurat_clusters))
  colnames(clu_df) <- c("Cluster", "Cells")
  clu_df$Pct <- paste0(round(100 * clu_df$Cells / sum(clu_df$Cells), 1), " %")
  .nice_dt(clu_df, caption = "Cluster sizes", page_len = 20)
}
```

### Cell proportions by condition

```{r proportion-plot, fig.height=5, fig.width=9}
if (.col_exists(obj, "seurat_clusters") && .col_exists(obj, "condition")) {
  prop_df <- obj@meta.data |>
    dplyr::count(condition, seurat_clusters) |>
    dplyr::group_by(condition) |>
    dplyr::mutate(pct = 100 * n / sum(n)) |>
    dplyr::ungroup()

  p <- ggplot(prop_df,
              aes(x = condition, y = pct,
                  fill = factor(seurat_clusters),
                  text = paste0("Cluster: ", seurat_clusters,
                                "<br>Condition: ", condition,
                                "<br>Proportion: ", round(pct, 1), " %"))) +
    geom_col(position = "stack") +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +
    labs(x = NULL, y = "Cell proportion (%)", fill = "Cluster") +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
  if (has_plotly) plotly::ggplotly(p, tooltip = "text") else p
}
```

---

## 4 · Cell Type Annotation {.tabset}

```{r singler-umaps, results='asis', fig.height=6, fig.width=9}
singler_cols <- grep("^singleR_labels", colnames(obj@meta.data), value = TRUE)

if (length(singler_cols) == 0L) {
  cat("> No SingleR annotation columns found.\n")
} else {
  for (sc in singler_cols) {
    label_pretty <- gsub("singleR_labels_", "SingleR ", sc)
    cat("\n### UMAP –", label_pretty, "\n\n")
    p_sr <- .umap_plotly(obj, sc,
                         title = paste("UMAP –", comp_group, "|", label_pretty),
                         point_size = 1.5)
    if (!is.null(p_sr)) print(p_sr)
    cat("\n")

    # annotation breakdown table
    cat("\n#### Annotation breakdown\n\n")
    ann_df <- as.data.frame(table(CellType = obj@meta.data[[sc]]))
    colnames(ann_df) <- c("Cell Type", "Cells")
    ann_df$Pct <- paste0(round(100 * ann_df$Cells / sum(ann_df$Cells), 1), " %")
    ann_df <- dplyr::arrange(ann_df, dplyr::desc(Cells))
    print(.nice_dt(ann_df,
                   caption = paste("Cell type counts –", label_pretty),
                   page_len = 15))
    cat("\n")
  }
}
```

---

## 5 · Differential Expression {.tabset}

```{r deg-section, results='asis'}
# Search only the cell-level DEG folder, excluding pseudobulk outputs.
deg_pattern <- "^(AllMarkers|DEG_Counts_).*\\.txt$"
deg_roots   <- list.dirs(out_dir, recursive = TRUE, full.names = TRUE)
deg_roots   <- deg_roots[basename(deg_roots) == "DEG"]
deg_files   <- .current_files(unlist(lapply(deg_roots, list.files, pattern = deg_pattern,
                             full.names = TRUE), use.names = FALSE))
deg_files   <- deg_files[
  !grepl("[\\\\/]Pseudobulk[\\\\/]", deg_files)
]

if (length(deg_files) == 0L) {
  if (!is.null(cfg$deg$run) && !isTRUE(cfg$deg$run)) {
    cat("> Cell-level DEG was disabled (`deg$run = FALSE`). Descriptive results may still be available in Expression, Proportions, Pseudobulk, Correlation, and GSEA folders.\n\n")
  } else {
    cat("> No cell-level DEG tables are available. This can occur when comparisons had fewer than two conditions, groups had too few cells, or no test results were produced.\n\n")
  }
} else {
  for (deg_f in deg_files) {
    tab_label <- gsub("AllMarkers_|_|.txt$", " ", basename(deg_f))
    cat("\n### ", trimws(tab_label), "\n\n")

    deg_tbl <- tryCatch(
      utils::read.table(deg_f, header = TRUE, sep = "\t",
                        stringsAsFactors = FALSE, check.names = FALSE),
      error = function(e) NULL
    )
    if (is.null(deg_tbl) || nrow(deg_tbl) == 0L) {
      cat("> Empty or unreadable file.\n\n")
      next
    }

    if (grepl("^DEG_Counts_", basename(deg_f))) {
      print(.nice_dt(deg_tbl, caption = basename(deg_f), page_len = 20))
      cat("\n")
      next
    }

    # Round numeric columns for readability
    num_cols <- vapply(deg_tbl, is.numeric, logical(1))
    deg_tbl[num_cols] <- lapply(deg_tbl[num_cols], round, digits = 4)

    print(.nice_dt(deg_tbl,
                   caption  = paste("Cell-level DEG –", trimws(tab_label)),
                   page_len = 20))
    cat("\n")
  }
}
```

### Average expression

```{r average-expression, results='asis'}
expr_files <- .current_files(list.files(out_dir, pattern = "^AvgExpr_.*\\.txt$",
                         recursive = TRUE, full.names = TRUE))
if (length(expr_files) == 0L) {
  if (!is.null(cfg$expression$layers) && length(cfg$expression$layers) == 0L) {
    cat("> Average-expression export was disabled by configuration.\n\n")
  } else {
    cat("> Average-expression tables are not available. Check whether configured assay layers were available.\n\n")
  }
} else {
  for (f in expr_files) {
    tab <- tryCatch(utils::read.delim(f, check.names = FALSE), error = function(e) NULL)
    cat("\n#### ", basename(f), "\n\n", sep = "")
    if (is.null(tab) || nrow(tab) == 0L) {
      cat("> Table unavailable or empty.\n")
    } else {
      print(.nice_dt(tab, caption = basename(f), page_len = 20))
    }
  }
}
```

### Proportion summaries

```{r proportion-tables, results='asis'}
proportion_files <- .current_files(list.files(out_dir, pattern = "^Proportions_.*\\.txt$",
                               recursive = TRUE, full.names = TRUE))
if (length(proportion_files) == 0L) {
  if (!is.null(cfg$proportions$run) && !isTRUE(cfg$proportions$run)) {
    cat("> Proportion analysis was disabled.\n\n")
  } else {
    cat("> Proportion tables are not available; two or more observed groups and conditions are required. See pipeline.log for details.\n\n")
  }
} else {
  for (f in proportion_files) {
    tab <- tryCatch(utils::read.delim(f, check.names = FALSE), error = function(e) NULL)
    cat("\n#### ", basename(f), "\n\n", sep = "")
    if (is.null(tab) || nrow(tab) == 0L) cat("> Table unavailable or empty.\n") else
      print(.nice_dt(tab, caption = basename(f), page_len = 20))
  }
}
```

### Replicate-aware pseudobulk DEG

```{r pseudobulk-deg, results='asis'}
pb_files <- .current_files(list.files(out_dir, pattern = "^PseudobulkDEG_.*\\.txt$",
                       recursive = TRUE, full.names = TRUE))
if (length(pb_files) == 0L) {
  if (!is.null(cfg$pseudobulk$run) && isTRUE(cfg$pseudobulk$run)) {
    cat("> Pseudobulk DEG was requested, but no results were produced. Check RNA counts, non-empty group/sample combinations, two conditions, replicate counts, and edgeR; see pipeline.log for details.\n\n")
  } else {
    cat("> Replicate-aware pseudobulk DEG was not enabled. Existing cell-level DEG uses cells as observations and is reported separately above.\n\n")
  }
} else {
  for (f in pb_files) {
    tab <- tryCatch(utils::read.delim(f, check.names = FALSE), error = function(e) NULL)
    if (is.null(tab) || nrow(tab) == 0L) {
      cat("> Pseudobulk table unavailable or empty: `", basename(f), "`.\n", sep = "")
    } else {
      cat("\n#### ", basename(f), "\n\n", sep = "")
      print(.nice_dt(tab, caption = basename(f), page_len = 20))
      cat("\n")
    }
  }
}
```

### Pseudobulk count exports

```{r pseudobulk-counts, results='asis'}
pb_count_files <- .current_files(list.files(out_dir, pattern = "^PseudobulkCounts_.*\\.txt$",
                             recursive = TRUE, full.names = TRUE))
if (length(pb_count_files) == 0L) {
  if (!is.null(cfg$pseudobulk$save_counts) && !isTRUE(cfg$pseudobulk$save_counts)) {
    cat("> Pseudobulk count export was disabled.\n\n")
  } else {
    cat("> Pseudobulk count export is not available. Check RNA counts, sample metadata, and replicate threshold; see pipeline.log.\n\n")
  }
} else {
  for (f in pb_count_files) {
    tab <- tryCatch(utils::read.delim(f, check.names = FALSE), error = function(e) NULL)
    cat("\n#### ", basename(f), "\n\n", sep = "")
    if (is.null(tab) || nrow(tab) == 0L) cat("> Table unavailable or empty.\n") else
      print(.nice_dt(utils::head(tab, 100L), caption = paste(basename(f), "(first 100 rows)"), page_len = 20))
  }
}
```

### Ranked GSEA

```{r gsea-results, results='asis'}
gsea_files <- .current_files(list.files(out_dir, pattern = "^GSEA_.*\\.tsv$",
                         recursive = TRUE, full.names = TRUE))
if (length(gsea_files) == 0L) {
  if (!is.null(cfg$gsea$run) && !isTRUE(cfg$gsea$run)) {
    cat("> Ranked GSEA was disabled in configuration.\n\n")
  } else {
    collection <- cfg$gsea$collection %||% "H"
    cat("> No ranked GSEA tables for MSigDB collection '", collection,
        "' are available. GSEA requires edgeR pseudobulk ranks, valid biological replicates in two conditions, and gene sets. Review pipeline.log for the specific skip reason.\n\n", sep = "")
  }
} else {
  for (f in gsea_files) {
    tab <- tryCatch(utils::read.delim(f, check.names = FALSE), error = function(e) NULL)
    if (is.null(tab) || nrow(tab) == 0L) {
      cat("> GSEA table unavailable or empty: `", basename(f), "`.\n", sep = "")
    } else {
      cat("\n#### ", basename(f), "\n\n", sep = "")
      print(.nice_dt(tab, caption = basename(f), page_len = 20))
      cat("\n")
    }
  }
}
```

### Gene correlations

```{r correlation-results, results='asis'}
cor_files <- .current_files(list.files(
  out_dir, pattern = "^Correlations_.*\\.(tsv|txt)$",
  recursive = TRUE, full.names = TRUE
))
if (length(cor_files) == 0L) {
  corr_x <- cfg$genes$corr_genes_x %||% NULL
  corr_y <- cfg$genes$corr_genes_y %||% NULL
  if (is.null(corr_x) || is.null(corr_y) || !length(corr_x) || !length(corr_y)) {
    cat("> Correlation analysis was not configured with both gene sets.\n\n")
  } else {
    cat("> No correlation tables were produced. Check whether genes were present in RNA data and whether groups met the 30-cell threshold; pipeline.log contains specific skip details.\n\n")
  }
} else {
  for (f in cor_files) {
    tab <- tryCatch(utils::read.delim(f, check.names = FALSE), error = function(e) NULL)
    cat("\n#### ", basename(f), "\n\n", sep = "")
    if (is.null(tab) || nrow(tab) == 0L) cat("> Correlation table unavailable or empty.\n") else
      print(.nice_dt(tab, caption = basename(f), page_len = 20))
  }
}
```

---

## 6 · Sex & Cell Cycle {.tabset}

### Sex scoring

```{r sex-umap, fig.height=5, fig.width=9}
sex_col <- "Assigned_Sex"
# BUG FIX: was "sex_assignment", a column run_sex_scoring() never creates (it
# writes "Assigned_Sex", see sex_scoring.R) -- this section reported "Sex
# scoring was not run" on every dataset, even when it had. qc_report.Rmd
# already looked for "Assigned_Sex" correctly; this file had drifted.
if (.col_exists(obj, sex_col)) {
  sex_umap <- .umap_plotly(obj, sex_col,
                            title = paste("UMAP –", comp_group, "| Sex"))
  if (!is.null(sex_umap)) sex_umap

  sex_df <- as.data.frame(table(Sex = obj[[sex_col]]))
  colnames(sex_df) <- c("Sex", "Cells")
  knitr::kable(sex_df, align = c("l", "r"))
} else {
  cat("> Sex scoring was not run or 'Assigned_Sex' column not found.")
}
```

### Cell cycle

```{r cycle-umap, fig.height=5, fig.width=9}
phase_col <- "Phase"
if (.col_exists(obj, phase_col)) {
  cycle_umap <- .umap_plotly(obj, phase_col,
                              title = paste("UMAP –", comp_group, "| Cell Cycle Phase"))
  if (!is.null(cycle_umap)) cycle_umap

  phase_df <- as.data.frame(table(Phase = obj[[phase_col]]))
  colnames(phase_df) <- c("Phase", "Cells")
  knitr::kable(phase_df, align = c("l", "r"))
} else {
  cat("> Cell cycle scoring was not run or 'Phase' column not found.")
}
```

---

## 7 · Additional Groupings {.tabset}

```{r extra-groupings, results='asis', fig.height=6, fig.width=9}
extra_cols <- setdiff(groupings, c("seurat_clusters",
                                    grep("^singleR", groupings, value = TRUE)))

if (length(extra_cols) == 0L) {
  cat("> No additional grouping columns to display.\n")
} else {
  for (ec in extra_cols) {
    if (!.col_exists(obj, ec)) next
    label_pretty <- gsub("_", " ", ec)
    cat("\n### UMAP –", label_pretty, "\n\n")
    p_ec <- .umap_plotly(obj, ec,
                         title = paste("UMAP –", comp_group, "|", label_pretty))
    if (!is.null(p_ec)) print(p_ec)
    cat("\n")
  }
}
```

---

## 8 · Per-cell pathway scoring {.tabset}

```{r escape-results, results='asis'}
escape_assays <- grep("^escape\\.", names(obj@assays), value = TRUE)
if (!length(escape_assays)) {
  if (!is.null(cfg$escape$run) && !isTRUE(cfg$escape$run)) {
    cat("> Per-cell pathway scoring was disabled (`escape$run = FALSE`).\n\n")
  } else {
    cat("> No escape assay is available. Check the optional escape/msigdb dependencies, gene-set availability, and pipeline.log.\n\n")
  }
} else {
  cat("Available per-cell pathway-score assays: ",
      paste(escape_assays, collapse = ", "), ".\n\n", sep = "")
}
```

---

## Pipeline diagnostics

```{r diagnostics, results='asis'}
if (!is.null(params$log_path) && file.exists(params$log_path)) {
  log_lines <- readLines(params$log_path, warn = FALSE)
  diagnostics <- grep("\\[(WARNING|ERROR)\\]", log_lines, value = TRUE)
  skips <- grep("\\[SKIP", log_lines, value = TRUE)
  if (length(diagnostics) || length(skips)) {
    diagnostic_df <- data.frame(
      Type = c(rep("Warning/error", length(diagnostics)),
               rep("Skipped/unavailable", length(skips))),
      Message = c(diagnostics, skips),
      stringsAsFactors = FALSE
    )
    print(.nice_dt(diagnostic_df, caption = "Pipeline diagnostics", page_len = 20))
  } else {
    cat("> No warnings, errors, or skipped analyses were logged.\n")
  }
} else {
  cat("> Pipeline log is not available.\n")
}
```

## 9 · Session Info {.tabset}

<details>
<summary>Click to expand</summary>

```{r session-info}
si        <- utils::sessionInfo()
pkg_df    <- do.call(rbind, lapply(si$otherPkgs, function(p)
  data.frame(Package = p$Package, Version = p$Version, stringsAsFactors = FALSE)))
pkg_df    <- pkg_df[order(pkg_df$Package), ]

knitr::kable(pkg_df, row.names = FALSE,
             caption = "Loaded packages at report generation time")
```

```{r r-version}
cat("R version:", R.version$version.string, "\n")
cat("Platform: ", R.version$platform,        "\n")
cat("Date:     ", format(Sys.time()),         "\n")
```

</details>
